Concurrent Programming in Harmony

### 2022-01-09 book/paper.tex
str implemented

{\small
\begin{tabular}{|l|l|}
\hline
Address & compute address from two components \\
\hline
Apply & pop <{m}> and <{i}> and apply <{i}> to <{m}>, pushing a value \\
\hline
Assert, Assert2 & pop <{b}> and check that it is <{True}>.  Assert2 also pops value to print \\
\hline
AtomicInc/Dec & increment/decrement the atomic counter of this context \\
\hline
Continue & no-op (but causes a context switch) \\
\hline
Choose & choose an element from the set on top of the stack \\
\hline
Cut & cut a set into its smallest element and the remainder \\
\hline
Del [<{v}>] & delete shared variable <{v}> \\
\hline
DelVar [<{v}>] & delete thread variable <{v}> \\
\hline
Dup & duplicate the top element of the stack \\
\hline
Frame <{m a}> & start method <{m}> with arguments <{a}>,
initializing variables.  \\
\hline
Go & pop context and value, push value on context's stack, and add to context bag \\
\hline
IncVar <{v}> & increment thread variable <{v}> \\
\hline
Jump <{p}> & set program counter to <{p}> \\
\hline
JumpCond <{e p}> & pop expression and, if equal to <{e}>, set program counter to <{p}> \\
\hline
Load [<{v}>] & push the value of a shared variable onto the stack \\
\hline
LoadVar [<{v}>] & push the value of a thread variable onto the stack \\
\hline
Move <{i}> & move stack element at offset <{i}> to top of the stack \\
\hline
$n$-ary <{op}> & apply $n$-ary operator <{op}> to the top $n$ elements on the stack \\
\hline
Pop & pop a value of the stack and discard it \\
\hline
Print & pop a value and add to the print history \\
\hline
Push <{c}> & push constant <{c}> onto the stack \\
\hline
ReadonlyInc/Dec & increment/decrement the read-only counter of this context \\
\hline
Return & pop return address, push \texttt{result}, and restore program counter \\
\hline
Sequential & pop an address of a variable that has sequential consistency \\
\hline
SetIntLevel & pop <{e}>, set interrupt level to <{e}>, and push old interrupt level \\
\hline
Spawn [eternal] & pop initial thread-local state, argument, and method and spawn a new context \\
\hline
Split & pop tuple and push its elements \\
\hline
Stop [<{v}>] & save context into shared variable <{v}> and remove from context bag \\
\hline
Store [<{v}>] & pop a value from the stack and store it in a shared variable \\
\hline
StoreVar [<{v}>] & pop a value from the stack and store it in a thread variable \\
\hline
Trap & pop interrupt argument and method \\
\hline
\end{tabular}
}

@@

>>>>

{\small
\begin{tabular}{|l|l|}
\hline
Address & compute address from two components \\
\hline
Apply & pop <{m}> and <{i}> and apply <{i}> to <{m}>, pushing a value \\
\hline
Assert, Assert2 & pop <{b}> and check that it is <{True}>.  Assert2 also pops value to print \\
\hline
AtomicInc/Dec & increment/decrement the atomic counter of this context \\
\hline
Continue & no-op (but causes a context switch) \\
\hline
Choose & choose an element from the set on top of the stack \\
\hline
Cut & cut a set into its smallest element and the remainder \\
\hline
Del [<{v}>] & delete shared variable <{v}> \\
\hline
DelVar [<{v}>] & delete thread variable <{v}> \\
\hline
Dup & duplicate the top element of the stack \\
\hline
Frame <{m a}> & start method <{m}> with arguments <{a}>,
initializing variables.  \\
\hline
Go & pop context and value, push value on context's stack, and add to context bag \\
\hline
IncVar <{v}> & increment thread variable <{v}> \\
\hline
Invariant \textit{end} & code for invariant follows.  Skip to \textit{end} + 1\\
\hline
Jump <{p}> & set program counter to <{p}> \\
\hline
JumpCond <{e p}> & pop expression and, if equal to <{e}>, set program counter to <{p}> \\
\hline
Load [<{v}>] & push the value of a shared variable onto the stack \\
\hline
LoadVar [<{v}>] & push the value of a thread variable onto the stack \\
\hline
Move <{i}> & move stack element at offset <{i}> to top of the stack \\
\hline
$n$-ary <{op}> & apply $n$-ary operator <{op}> to the top $n$ elements on the stack \\
\hline
Pop & pop a value of the stack and discard it \\
\hline
Print & pop a value and add to the print history \\
\hline
Push <{c}> & push constant <{c}> onto the stack \\
\hline
ReadonlyInc/Dec & increment/decrement the read-only counter of this context \\
\hline
Return & pop return address, push \texttt{result}, and restore program counter \\
\hline
Sequential & pop an address of a variable that has sequential consistency \\
\hline
SetIntLevel & pop <{e}>, set interrupt level to <{e}>, and push old interrupt level \\
\hline
Spawn [eternal] & pop initial thread-local state, argument, and method and spawn a new context \\
\hline
Split & pop tuple and push its elements \\
\hline
Stop [<{v}>] & save context into shared variable <{v}> and remove from context bag \\
\hline
Store [<{v}>] & pop a value from the stack and store it in a shared variable \\
\hline
StoreVar [<{v}>] & pop a value from the stack and store it in a thread variable \\
\hline
Trap & pop interrupt argument and method \\
\hline
\end{tabular}
}

@@

\section{Formal Specification}
\label{ap:formal}

A formal specification of the Harmony Virtual Machine is well underway but
not yet completed.  At this time, there is no support yet for bitwise integer
operations.
Also, strings are limited to the printable characters minus
double quotes, back quotes, or backslashes.
Given a Harmony program, you can output the TLA+ specification for
the program using the following command:

\begin{code}
\begin{verbatim}
$ harmony -o program.tla program.hny
\end{verbatim}
\end{code}

For most Harmony programs, including Peterson's algorithm and the Dining
Philosophers in this book, the result is complete enough to run through
the TLC model checker.
